---
title: "DATA 304/COMP312 - Group 2 Project"
author: "Group 2"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
  rmarkdown::pdf_document:
    toc: yes
    number_sections: yes
    keep_md: yes
    fig_caption: yes
mainfont: Times New Roman
header-includes:
- \usepackage[T1]{fontenc}
- \renewcommand{\topfraction}{.85}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r results=FALSE, echo=FALSE}
library(pander)
```

# Report 2 - Submit Collected Data

## Group Members

```{r, fig.cap="table of group 4 members", echo=FALSE}

data <- data.frame(Names = c("Jacob French", "Jayden Hooper", "Reece Pacia", "Alif Nurrokhman", "Aston Robbins","James Boyle"),
                   Emails = c("frenchjaco@vuw.ac.nz","hooperjayd@myvuw.ac.nz","paciareec@myvuw.ac.nz", "nurrokmuha@my.vuw.ac.nz","robbinasto@myvuw.ac.nz","boylejame1@myvuw.ac.nz"),
                   ORCID = c('0009-0004-7268-3581', "0009-0004-2644-5216","0009-0001-4748-604X", "0009-0008-7814-9231", "0009-0005-2959-883X","0009-0002-1177-2922" ))
pander(data)                  
```

## Second Meeting w/ Service Identification

On the 19th of April we had 2 new members added to our group Jacob
French and Alif Nurrokhman as their previous group had disbanded. We had
a meeting with Jayden, Jacob and Alif. We further discussed the WCT
idea, however we discovered that although we could access individual
waiting and service times for the queue we had not been given access to
the individual inter-arrival times. Jayden emailed his manager to
inquire about this. Jacob and Alif also brought up that they had
previously gained permission to observe the Louis Cafe queue in their
previous group.

We had another group meeting on the 22nd of April with all members of
our group. In this meeting Jayden confirmed that unfortunately his
manager could not give us access to the inter-arrival times. As we could
no longer gather the correct data from WCT we decided against this idea.
Instead we decided to gather data from the louis cafe.

## Identifying Tools

The tools we decided to use for our investigation were:

-   Notebooks, pens, and laptops for recording and capturing data (one
    per group member)
-   Jupyter notebook to model the queues in Python (specifically using
    the SimPy, NumPy, and Panda packages)
-   Discord for communication purposes
-   GitLab for creating a Git repository
-   RMarkdown to write the report
-   One person to track the queue, one to time the arrivals, and one to
    time the service
-   Online stopwatch in order to time multiple people at once
-   Google sheets document with data set, to be later converted into CSV
    as a machine-readable format

## Louis Cafe Data

```{r}
cafe = read.csv("cafe Louis Data.csv")
pander(head(cafe,10))
```

## Plots

```{r,echo=FALSE}
library(ggplot2) ;library(gridExtra);library(dplyr)

a <- ggplot(cafe, aes(x=Inter.Arrival.Time)) + geom_histogram() + labs(x='Inter-Arrival Time (Seconds',y ="Count")

b <- ggplot(cafe, aes(x=Waiting.Time)) + geom_histogram() + labs(x='Waiting Time (seconds)',y ="Count")

c <- ggplot(cafe, aes(x=Service.Time)) + geom_histogram() + labs(x='Service Times (seconds)',y ="Count")

grid.arrange(a,b,c,nrow = 2)
```
