---
title: "Simulation and Stocahstic Models - Group 2 Project"
author: "A Study of Server Capacity at Cafe Louis"
date: "Jacob French - Alif Nurrokhman - Jayden Hooper - Reece Pacia - James Boyle"
output:
  pdf_document:
    df_print: paged
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Introduction

```{r results=FALSE, echo=FALSE}
library(pander)
```

## Members

```{r, fig.cap="table of group 4 members", echo=FALSE}

data <- data.frame(Names = c("Jacob French", "Jayden Hooper", "Reece Pacia", "Alif Nurrokhman","James Boyle"),
                   Emails = c("frenchjaco@myvuw.ac.nz","hooperjayd@myvuw.ac.nz","paciareec@myvuw.ac.nz", "nurrokmuha@myvuw.ac.nz","boylejame1@myvuw.ac.nz"),
                   ORCID = c('0009-0004-7268-3581', "0009-0004-2644-5216","0009-0001-4748-604X", "0009-0008-7814-9231","0009-0002-1177-2922" ))
pander(data)
```

## Terminology

-   Arrival Rate - $\lambda$
-   Service Rate - $\mu$
-   Steady-State Attainable if - $\rho = \frac{\lambda}{\mu} < 1$
-   Number of Servers - c
-   Expected Waiting Time in System - W

## Introduction

Queuing is traditionally defined as when people, materials or
information must wait at certain times to receive a service (Teknomo, K
2012). This study will explain the basics of queuing theory and its
relation to organisational performance and customer satisfaction in the
hospitality industry. Queuing theory is frequently used to assess
customer waiting time, server utilisation, earnings and costs. In
hospitality, companies service's often cannot keep up with their
arrivals, which results in long waiting times and loss of potential
profits. In this study, we will study Cafe Louis's queuing system and
prepare theoretical and simulation models that could be used to predict
what effects changes in queue setup, such as changes in load, would
have.

## Problem Statement

Businesses have poor organisational performance and customer
satisfaction due to service not keeping up with arrivals. If queuing
theory was applied this could be maximised.

## Objectives

-   Examine whether a change in queue setup will increase organisational
    performance, to be measured using profit.
-   Examine whether a change in queue setup will increase customer
    satisfaction, to be measured using waiting time.

## Hypothesis

Based on the objectives of our study, the following hypothesis will be
tested.

-   Will increasing server capacity increase customer satisfaction and
    organisational performance?

## Expectations

-   Steady-State distribution is not attainable under current
    conditions. This will cause waiting time to tend to infinity.
-   Organisation performance is not being maximised as service cannot
    keep up with arrivals, resulting in a loss of sales.
-   Customer Satisfaction is not being maximised as service cannot keep
    up with arrivals, which results in long waiting times.
-   Increasing Server capacity will fix all of the above.

## Literature Review

For this study, we followed a handful of similar research papers done in
this field in which they observed real-world systems and the effect that
waiting time has on customer satisfaction. Queuing becomes an
inconvenience and can be highly costly for organisations. Problematic
queuing systems can significantly harm the customer's overall
satisfaction (Ronald Anthony Nosek, Jr., & James P. Wilson, 2001).

A highly satisfied customer will likely provide repeat business,
increasing revenues and profitability (Ronald Anthony Nosek, Jr., &
James P. Wilson, 2001). Therefore we wanted the focus of this study to
determine what factor could be optimised for the overall performance of
an organisation to increase customer satisfaction. There are two
approaches to improving customer satisfaction with waiting time: through
decreasing actual waiting time as well as through enhancing customer
waiting experience (Katz, Larson, and Larson, 1991; Davis and Heineke,
1994)

Because we cannot influence the waiting experience, a key metric we will
use in this study is overall waiting time, as the amount of time
customers must spend waiting can significantly affect their satisfaction
(Davis and Vollman, 1990). Ronald Anthony Nosek, Jr., & James P. Wilson
(2001) also considers that the measurement of customer satisfaction --
is highly qualitative and subjective, and the relationship is generally
inverse (i.e., in general, as waiting time decreases, satisfaction
increases).

Teknomo, K (2012) states that a queuing system has two primary
components: the customers and servers. Teknomo also says in their study
on construction management that there is a direct relationship between
the server and the customer. They stated that a system with fewer
servers could incur customer costs (satisfaction), and adding servers to
the system can incur higher server costs (organisation). As we think in
terms of [the entire] system, the total cost of [the] queuing system
must include both the server side and the customer side (Teknomo, K,
2012). Therefore, the optimum number of servers is the one that
minimises the total cost (Teknomo, K, 2012).

For the system we observed for this study, we state that the time it
takes for a server to complete a request strongly influences the overall
waiting time of a customer. We can optimise waiting time by optimising
the server side of the system (hence our hypothesis). For an
organisation to perform optimally, customer sanctification and
operational costs are maximised. Because queuing theory is rich in
optimisation (Teknomo K, 2012), we will utilise the tools that queuing
theory provides to attain precise performance measures to assess
real-world systems.

# Methodology

## Data Collection

Before collecting data, we obtained permission to do so via email. Once
permission was granted, we began compiling our data between 11:00-14:00
on weekdays, as this window was peak busy hours for the cafe, and we
could collect data much faster. We collected data on the system's metric
arrival times, service times and waiting times. We ensured that our
data-collecting methods were not invasive, including not collecting
personal information or disrupting the queue itself.

The tools we decided to use for our data collection were:

-   Notebooks, pens, and laptops for recording and capturing data (one
    per group member)
-   Online stopwatch to time multiple people at once
-   Google Sheets document with the data set, to be later converted into
    CSV as a machine-readable format

## Data Analyse Methods

For analysing the data, we used Python for exploratory analysis and
utilised the 'disfit' package to find and plot the "best-fit"
distribution for our data. Additionally, we used a Python function
provided by Dr Binh to perform chi-square goodness of fit tests. Other
Python packages we used were pandas, numpy, matplotlib and scipy.

# Reproducible Repository

We set up a reproducible repository described by Frery, Gomez & Medeiros
(2020) on GitHub so that this study may be reproduced.
[312-Group-2-Project](https://github.com/AleafNurr/312-Group-2-Project){.uri}

# Observed Data

## Arrival, Service and Waiting Times

```{r, out.width="50%",echo=FALSE}
knitr::include_graphics(c("Inter-Arrivals.pdf","Waiting_time.pdf","Service_time.pdf"))
```

## Exploratory Statistics

```{r, echo=FALSE}
obs_stat_table = data.frame(
  Metric = c("Count", "mean", "std", "min", "25%", "50%", "75%", "max"),
  Inter_ArrivalTime = c(251, 27.67, 33.07, 0.00, 6.52, 17.37, 34.33, 229.52),
  WaitingTime = c(251, 66.52, 59.55, 0.00, 18.27, 56.61, 90.40, 247.60),
  ServiceTime = c(251, 41.30, 28.10, 5.97, 23.73, 33.5, 49.35, 239.34)
)
pander(obs_stat_table, caption = "Observed Data")
```

```{r, echo=FALSE}
exp_stat_table = data.frame(
  Metric = c("Expected Inter-Arrival Times", "Expected Service Time", "Waiting Time"),
  "Mean(s)" = c("27.7", "41.3", "66.5"),
  "Rates per Hour" = c("130", "87", "--")
)
pander(exp_stat_table, caption = "Expected Metrics based off Observed Data")
```

Based on this table, we find that the rate of customers arriving is more
significant than that of customers being served such that $\lambda>\mu$,
therefore, steady-state is unattainable for the current system.

# Models

We built three models to represent our system. - Model 1 uses
exponential distribution for arrival times and service times - Model 2
uses exponential for arrivals and Erlang for service - Model 3 uses
exponential distribution for arrival times and uses empirical
distribution for service times

## Model 2

## Best Fit Distributions

### Inter-Arrival Times

Using the distfit package, we determined that the most suitable
probability distribution for modelling inter-arrival times is the
Exponential distribution, compared to the Erlang and Gamma
distributions.

```{r, out.width="100%",echo=FALSE}
knitr::include_graphics("best1.pdf")
```

## Chi-Square GOF

$$
H_0: \text{The sample is exponentially distributed}
$$

$$
H_1: \text{The sample is not exponentially distributed}
$$

However, we conducted a chi-squared goodness-of-fit test following the
null hypothesis that the sample data follows an Exponential distribution
and the alternative hypothesis that it does not. The test resulted in a
chi-square statistic of 64.836 with a corresponding p-value of
2.184e-05. These results suggest insufficient evidence to support the
null hypothesis that the sample is exponentially distributed. Therefore,
we cannot conclude that the inter-arrival times are Exponentially
distributed. Further investigation may be necessary to identify a more
appropriate distribution for modelling the inter-arrival times.

## Service Times

Using the distfit package, we identified the most suitable probability
distribution for modelling service time as the Erlang distribution after
comparing it to the Exponential and Gamma distributions.

```{r, out.width="100%",echo=FALSE}
knitr::include_graphics("best2.pdf")
```

Erlang.$$
H_0: \text{The sample is Erlang distributed}
$$

$$
H_1: \text{The sample is not Erlang distributed.}
$$

We then conducted a chi-squared goodness-of-fit test with the null
hypothesis that the sample data follows an Erlang distribution and the
alternative hypothesis that it does not. The resulting chi-square
statistic was 26.492 with a corresponding p-value of 0.546. Therefore we
conclude that there is sufficient evidence to accept the null hypothesis
and conclude that the service times are Erlang distributed. This implies
that the Erlang distribution provides a good fit for modelling the
service times in our study.

## Model Results

Simulating the models for three hours produced the following results:

```{r, echo=FALSE}
exp_stat_table <- data.frame(
  Model = c("Model 1:M/M/1", "Model 2:Best-fit Distribution", "Model 3:Empirical Distribution"),
  "Mean Waiting Time(s)" = c(1872.87, 1275.37, 1704.03)
)
pander(exp_stat_table, caption = "Model Results",digits = 6)
```

Because a steady state is not attainable, all these results will
approach infinite. Once we add more server capacity, steady-state should
be achievable, and we expect that model 2 will be the best
representation of the system.

The reason why we would not expect model 1 to be the best representation
is possibly underfitting, as the distribution used for this model may
not be the best representation.

We would not expect model 3 to be the representation because the model
may be overfitting when using the empirical distribution.

Model 2 produces the shortest waiting time, and because it is fitted
with the 'best fit' distribution, it should be the most accurate
representation of our model. We will continue to analyse Model 2.

# Cost Analysis

## Assumptions

For this section, we make the following assumptions:

-   Servers are paid a minimum wage of \$22.70 per hour.
-   Server cost per hour = c \* 22.70
-   The average spend = \$5 (estimated as the item's average cost on
    men. u)Earnings per hour = average spend \* service rate per hour.
    Whilst serving rate \< arrival rate. Otherwise, service rate per
    hour = arrival rate.
-   Earnings lost per hour = (service rate per hour - arrival rate per
    hour) \* average spend, so does this. Whilst Arrival rate \>=
    Service rate. Otherwise, earnings lost per hour = 0.
-   Profit = Earnings per hour - server cost per hour -Assume all other
    costs are covered.

## Results

Changing server capacity produces the following results:

```{r, echo=FALSE}
exp_stat_table <- data.frame(
  "Server Capacity (c)" = c("1", "2", "3", "4"),
  "Expected Waiting Time" = c(NA, 77.162, 45.96, 42.22),
  "Arrival Rate" = c(130, 130, 130, 130),
  "Service Rate" = c(87, 130, 130, 130),
  "Utilisation Factor" = c(1.49, 0.86, 0.79, 0.78),
  "Earnings per hour ($NZD)" = c(435.81, 654.3, 654.3, 654.3),
  "Earnings Lost per hour ($NZD)" = c(218.48, 0, 0, 0),
  "Profit per hour ($NZD)" = c(413.11, 608.9, 586.2, 564.3)
)

pander(exp_stat_table, caption = "Change in Capacity Results",digits = 6)
```

Results indicate that profit is maximised when the system has two
servers; if the system has two or more servers, gain follows a linear
trend 654.3 -wages \* c. Results also indicate that waiting time shows
no significant decrease after the system has three servers. This can be
observed graphically.

## Profit & Waiting Time Plot

```{r, out.width="100%",echo=FALSE}
knitr::include_graphics("tradeoff1.pdf")
```

## Profit Waiting Time Trade-off

As the optimal number of metrics servers disagree, we have a trade-off.
Hence we suggest a new metric profit-waiting time to find the number of
servers that balances profit and waiting time. This can be observed
graphically.

```{r, out.width="100%",echo=FALSE}
knitr::include_graphics("tradeoff.pdf")
```

The graph shows that profit - waiting time will be maximised if the
system has three servers.

# Conclusion & Discussion

## Recommendations

In conclusion, our models indicate Cafe Louis does the following between
11-2 pm.

-   If Cafe Louis values organisation performance over customer
    satisfaction, they should hire one additional staff member as this
    will maximise their profit.
-   If Cafe Louis values customer satisfaction over organisational
    performance, they should hire two or more staff members, as this
    will decrease their waiting times significantly.
-   If Cafe Louis wants to find a balance between the two, they should
    hire two staff members as this maximises the difference between
    profit and waiting times.

## Discussion

We believe the waiting times calculated by our models were inaccurate.
This is because Cafe Louis had busier periods between lectures(roughly).
During these busier periods, the majority of our data points were
collected. We believe this directly affected our waiting times and
arrival rates; hence, we didn't have constant arrivals (an exponential
distribution assumption). We believe the simulation models we built
better represent this period rather than 11-2 pm. Suppose data was only
observed and simulated for this period. In that case, we believe it
would fix problems such as small p-values for the chi-square goodness of
fit tests for exponentially distributed arrival rates, and we would see
an increase in waiting time when adding additional servers. If a study
was carried out on this, we believe it would produce significantly
better results.

## Evidence

Simulating our models for ten minutes rather than three hours results in
much closer waiting time estimates to our observed mean. This is
evidence that our models represent this period better.

```{r, echo=FALSE}
exp_stat_table <- data.frame(
  Model = c("Model 1:M/M/1", "Model 2:Best-fit Distribution", "Model 3:Empirical Distribution","Observed"),
  "Mean_Waiting_Time(s)_3hours" = c(1872.87, 1275.37, 1704.03,66),
  "Mean_Waiting_Time(s)_10mins" = c(186.3,11.75,205.72,66)
)
pander(exp_stat_table, caption = "Model Results",digits = 6)
```

## Conclusion

In conclusion, increasing server capacity will improve customer
satisfaction and organisational performance.

## References

-   Aina, M. O. Y. P. N. B., & Kazeem, O. Queuing Theory and Customer
    Satisfaction: A Review of Performance, Trends and Application in
    Banking Practice (A Study of First Bank Plc Gwagwalada, Abuja
    Branch).

-   [Davis,
    M.M.](https://www.emerald.com/insight/search?q=Mark%20M.%20Davis "Mark M. Davis")
    and [Heineke,
    J.](https://www.emerald.com/insight/search?q=Janelle%20Heineke "Janelle Heineke")
    (1994), "Understanding the Roles of the Customer and the Operation
    for Better Queue Management", [*International Journal of Operations
    & Production
    Management*](https://www.emerald.com/insight/publication/issn/0144-3577),
    Vol. 14 No. 5, pp. 21-34.
    <https://doi.org/10.1108/01443579410056777>

-   Frery, A. C., Gomez, L., & Medeiros, A. C. (2020). A badging system
    for reproducibility and replicability in remote sensing research.
    IEEE Journal of Selected Topics in Applied Earth Observations and
    Remote Sensing, 13, 4988-4995.

-   Katk. K, Larson. K, Larson. R. (1991). Prescription for Waiting Line
    Blues: Entertained, Enlighten and Engage, Sloan Management Review,
    <https://www.researchgate.net/publication/304582002_Prescription_for_the_Waiting_in_Line_Blues_Entertain_Enlighten_Engage>

-   Teknomo, K. (2012). Queuing rule of thumb based on M/M/s queuing
    theory with applications in construction management. Civil
    Engineering Dimension, 14(3), 139-146.

-   Ronald Anthony Nosek, Jr., & James P. Wilson. (2001). Queuing Theory
    and Customer Satisfaction: A Review of Terminology, Trends, and
    Applications to Pharmacy Practice.

-   <https://www.math.tamu.edu/~phoward/m442/queue.pdf>
